---
title: "Half Three Pizza"
author: "Syarif Yusuf Effendi"
output: github_document
---

# Summary

A year's worth of sales from a fictitious pizza place, including the date and time of each order and the pizzas served, with additional details on the type, size, quantity, price, and ingredients.

# ASK

#### Key objective:

-   How many customers do we have each day? Are there any peak hours?
-   How many pizzas are typically in an order? Do we have any bestsellers?
-   How much money did we make this year? Can we identify any seasonality in the sales?
-   Are there any pizzas we should take of the menu, or any promotions we could leverage?

# PREPARE

#### Key objective:

1.  Determine the credibility of data:

-   This data is obtained from the [Maven Analytics](https://www.mavenanalytics.io/data-playground) page. The data contains several tables related to sales, types, and other information. Each table has a relationship with each other

2.  Join the whole table into a dataset
3.  Install and load necessary packages:

-   `tidyverse`
-   `lubridate`
-   `dplyr`

```{r}
library(tidyverse)
```

```{r}
library(dplyr)
```

```{r}
library(lubridate)
```

4.  Import data to RStudio

```{r}
order_details <- read_csv("D:/Dataset/Pizza Sales/order_details.csv")
```

```{r}
orders <- read_csv("D:/Dataset/Pizza Sales/orders.csv")
```

```{r}
pizzas <- read_csv("D:/Dataset/Pizza Sales/pizzas.csv")
```

```{r}
pizza_types <- read_csv("D:/Dataset/Pizza Sales/pizza_types.csv")
```

5.  Merge all tables

```{r}
pizza.order <- inner_join(order_details, orders, by="order_id")
pizza.type <- inner_join(pizza_types, pizzas, by="pizza_type_id")
pizza.dataset <- inner_join(pizza.order, pizza.type, by="pizza_id")
```

```{r}
head(pizza.dataset)
```

# Process

#### Key objective:

1.  Clean and prepare the data for analysis

-   Now that all the data is in one location, we can begin to remove any mistakes like NA. In order to speed up our research and produce more insightful results, we will also make some adjustments to the data, adding relevant new columns based on calculations of previously existing columns

```{r}
glimpse(pizza.dataset)
```

```{r}
summary(pizza.dataset)
```

#### Create new columns

```{r}
pizza.dataset$weekday <- weekdays(pizza.dataset$date)
pizza.dataset$month <- month(pizza.dataset$date)
pizza.dataset$hour <- hour(pizza.dataset$time)
pizza.dataset$revenue <- pizza.dataset$quantity * pizza.dataset$price
```

```{r}
glimpse(pizza.dataset)
```

```{r}
summary(pizza.dataset)
```

# ANALYZE

We now have completed data frame with all of the information we need and then export the data frame into a csv file and perform analysis and visualization through Tableau Public

```{r}
write.csv(pizza.dataset, "D:/Dataset/Pizza Sales/pizza.dataset.csv", row.names = FALSE)
```

Click [here](https://public.tableau.com/views/PizzaDashboard_16691223876820/SalesDashboard?:language=en-US&:display_count=n&:origin=viz_share_link) to see the result of my Tableau Dashboard

# SHARE

#### Conclusion

1.  The average daily customers are 60 customers. There is a peak hours every lunchtime on weekdays, and dinnertime on weekends.
2.  Up to 67% of customers placed orders for 1-2 pizzas, and the remaining clients placed orders for 3 or more pizzas. A popular pizza is the Big Meat Pizza in small portions.
3.  The restaurant made \$817.8K in sales for the entire year. With Q2 having the highest average income.
4.  The Greek Pizza size XXL needs to be removed from the menu and an additional Brie Carre Pizza size option added.

#### Recommendation

1. Lunchtime on weekdays and dinnertime on weekends are peak sales periods. At these situations, the staff must prioritize preparation.
2. Alternate Operational Hours: 11.00 - 21.59. Potential to reduce the weight of variable costs while just 3%, or $24.3K, of annual revenue is lost.
3. Increasing the price of the top five pizzas or increasing the pizza size to XL on pizzas with a maximum size of L can be selected to boost revenue.
4. The Brie Carre pizza has the second greatest revenue in the small pizza group. It is important to provide a larger size option and offer a discount to promote the new size.
5. The Greek pizza sells better at the XL size, so the XXL version can be removed from the menu. The most popular pizzas can be made in XXL size as a replacement.
6. n raise the quantity of daily pizza sales to boost revenue in the upcoming year. Giving discounts on some items, particularly at breakfast and at midnight, might increase daily sales